Getting start

### 2023-11-21 inst/gitbook/01_getting_started.Rmd
fully updated manual

Et voila! We now have a distribution of 50 trees and 50 datasets from which we can calculate disparity.
We can convert this list into `dispRity` using the `treats.dispRity` function.

```{r, message = FALSE}
## Converting the data to be passed to dispRity functions
dispRity_data <- dispRitreats(sim_data)

## Creating the time series (by default dispRitreats scales the trees to be 1 unit of times old)
time_series <- chrono.subsets(data = dispRity_data$data,
                              tree = dispRity_data$tree,
                              time = seq(from = 0.9, to = 0, by = -0.1),
                              method = "continuous", model = "proximity")

## Calculating the disparity from the simulated data
simulated_disparity <- dispRity(time_series, metric = c(sum, variances),
                                verbose = TRUE)
par(mfrow = c(1,2))
plot(simulated_disparity, main = "simulated disparity")
plot(change_in_size, main = "observed disparity")
```

@@

```{r}
## Converting the data to be passed to dispRity functions
rand_ext_data <- dispRitreats(sim_rand_extinction)
trait_ext_data <- dispRitreats(sim_trait_extinction)

## Creating the time series
rand_ext_time_series <- chrono.subsets(data = rand_ext_data$data,
                                       tree = rand_ext_data$tree,
                                       time = seq(from = 0.9, to = 0, by = -0.1),
                                       method = "continuous", model = "proximity")
trait_ext_time_series <- chrono.subsets(data = trait_ext_data$data,
                                        tree = trait_ext_data$tree,
                                        time = seq(from = 0.9, to = 0, by = -0.1),
                                        method = "continuous", model = "proximity")

## Calculating the disparity from the simulated data
sim_rand_ext_disp <- dispRity(rand_ext_time_series, metric = c(sum, variances),
                              verbose = TRUE)
sim_trait_ext_disp <- dispRity(trait_ext_time_series, metric = c(sum, variances),
                               verbose = TRUE)

par(mfrow = c(2,2))
plot(change_in_size, main = "observed disparity")
plot(simulated_disparity, main = "simulated disparity")
plot(sim_rand_ext_disp, main = "simulated disparity\n(mass extinction)")
plot(simulated_disparity, main = "simulated disparity\n(trait extinction)")
```

Interestingly, with all our parameters used here and our specific disparity metric we cannot detcet the effect of a mass extinction.

>>>>

Et voila! We now have a distribution of 50 trees and 50 datasets from which we can calculate disparity.
We can directly pass the results to the `dispRity` package pipeline using the `dispRitreats` function:

```{r, message = FALSE, warning=FALSE}
## Calculate the dispRity for all the simulations
simulated_disparity <- dispRitreats(sim_data,
                                    method = "continuous",
                                    model  = "proximity",
                                    time   = 10,
                                    metric = c(sum, variances),
                                    scale.trees = TRUE)
par(mfrow = c(1,2))
plot(simulated_disparity, main = "simulated disparity")
plot(change_in_size, main = "observed disparity")
```

@@

```{r}
## Simulate the tree and traits with a random extinction event
sim_rand_extinction <- treats(
                   traits     = my_traits,
                   bd.params  = est_params,
                   stop.rule  = stop_rule,
                   events     = random_extinction,
                   null.error = 100,
                   replicates = 50)

Interestingly, with all our parameters used here and our specific disparity metric we cannot detect the effect of a mass extinction.
